---
title:  "MODEL_COMPARISON_FIGURE3"
author: "Matthew J Shannon"
date:   "18/01/2024"
output: html_document
---

# Load in all dependencies

```{r, include = FALSE}
devtools::install_github ('chris-mcginnis-ucsf/DoubletFinder')
devtools::install_github ('satijalab/seurat-wrappers')
remotes::install_github  ("mojaveazure/seurat-disk")
devtools::install_github ("velocyto-team/velocyto.R")
BiocManager::install     ("pcaMethods")
BiocManager::install     ('MAST')
BiocManager::install     ('org.Hs.eg.db')
BiocManager::install     (version = "3.12")
BiocManager::install     ('Seurat')
BiocManager::install     ('readxl')
BiocManager::install     ('modes')
BiocManager::install     ('pheatmap')
BiocManager::install     ('limma')
BiocManager::install     ('clustree')
BiocManager::install     ('clusterProfiler')
BiocManager::install     ('EnhancedVolcano')
install.packages         ('fields')
install.packages         ('plotly')
install.packages         ("VennDiagram")
library                  (Seurat)
library                  (readr)
library                  (tidyverse)
library                  (Matrix)
library                  (dplyr)
library                  (ggplot2)
library                  (cowplot)
library                  (tibble)
library                  (readxl)
library                  (sctransform)
library                  (fields)
library                  (KernSmooth)
library                  (ROCR)
library                  (parallel)
library                  (reshape2)
library                  (pheatmap)
library                  (DoubletFinder)
library                  (limma)
library                  (SeuratWrappers)
library                  (SeuratDisk)
library                  (plotly)
library                  (clustree)
library                  (velocyto.R)
library                  (MAST)
library                  (EnhancedVolcano)
library                  (clusterProfiler)
library                  (AnnotationDbi)
library                  (org.Hs.eg.db)
library                  (VennDiagram)
library                  (RColorBrewer)
library                  (EnhancedVolcano)
library                  (xlsx)
library                  (SCENIC)
library                  (AUCell)
library                  (GENIE3)
library                  (RcisTarget)
library                  (monocle3)
gc                       ()
```

# Load the integrated and model-specific trophoblast Seurat objects

```{r}
# Integrated (all models) trophoblast object
load         (file = "~/Desktop/FINAL_Trophoblast_Data.Rdata")
DefaultAssay (Trophoblasts) <- "integrated"

# MPI (in vivo) trophoblast object
load         (file = "~/Desktop/MPI.Rdata")
DefaultAssay (MPI) <- "integrated"

# TBpOrg (in vitro) trophoblast object
load         (file = "~/Desktop/TBpOrg.Rdata")
DefaultAssay (TBpOrg) <- "integrated"

# TBsOrg (in vitro) trophoblast object
load         (file = "~/Desktop/TBsOrg.Rdata")
DefaultAssay (TBsOrg) <- "integrated"
```

# Figure 3 and Supplemental Figure 3

## Integrated (in vivo + in vitro) CTBs

```{r}
CTBs <- subset (Trophoblasts, idents = c("CTB 1", "CTB 2", "CTB 3", "CTB 4"))
CTBs
```

```{r, Figure 3A, integrated CTBs}
UMAPPlot (CTBs, label = TRUE, pt.size = 3, cols = c("#E31A1C", "#999999", "#225EA8", "#984EA3")) + NoLegend ()
```

## Differential gene expression analysis

```{r, Figure 3B MPI vs TBpOrg differential gene expression analysis}
DefaultAssay (CTBs) <- "RNA"

TBpOrg.CTB.response <- FindMarkers (CTBs,
                                    ident.1         = c("MPI"),
                                    ident.2         = c("TBpOrg"),
                                    test.use        = "MAST",
                                    logfc.threshold = -Inf,
                                    min.pct         = -Inf,
                                    min.diff.pct    = -Inf,
                                    verbose         = FALSE)
                       head        (TBpOrg.CTB.response, n = 100)
```

```{r, Table S6A}
write.xlsx (TBpOrg.CTB.response, file = "MPI_TBpOrg_CTB_DGE.xlsx")
```

```{r, Figure 3B MPI vs TBpOrg volcano plot}
EnhancedVolcano (TBpOrg.CTB.response,
                 lab             = rownames (TBpOrg.CTB.response),
                 x               = "avg_log2FC",
                 y               = "p_val_adj",
                 FCcutoff        = 1.00,
                 pCutoff         = 10e-50,
                 pointSize       = 4.0,
                 col             = c('grey', 'grey', 'grey', 'red'),
                 colAlpha        = 1,
                 selectLab       = c('SMAGP', 'SMAGP', 'EGR1', 'BCAM', 'GPX3', 'SPINT1', 'CLU', 'BRI3', 'IFI27', 'ISG15', 'NEAT1'), 
                 xlim            = c(-4.1, 4.1),
                 gridlines.major = FALSE,
                 drawConnectors  = TRUE, 
                 gridlines.minor = FALSE) + NoLegend () + ggtitle (NULL)
```

```{r, Figure 3B MPI vs TBsOrg differential gene expression analysis}
DefaultAssay (CTBs) <- "RNA"

TBsOrg.CTB.response <- FindMarkers (CTBs,
                                    ident.1         = c("MPI"),
                                    ident.2         = c("TBsOrg"),
                                    test.use        = "MAST",
                                    logfc.threshold = -Inf,
                                    min.pct         = -Inf,
                                    min.diff.pct    = -Inf,
                                    verbose         = FALSE)
                       head        (TBsOrg.CTB.response, n = 100)
```

```{r, Table S6A}
write.xlsx (TBsOrg.CTB.response, file = "MPI_TBsOrg_CTB_DGE.xlsx")
```

```{r, Figure 3B MPI vs TBsOrg volcano plot}
EnhancedVolcano (TBsOrg.CTB.response,
                 lab             = rownames (TBsOrg.CTB.response),
                 x               = "avg_log2FC",
                 y               = "p_val_adj",
                 FCcutoff        = 1.00,
                 pCutoff         = 10e-50,
                 pointSize       = 4.0,
                 col             = c('grey', 'grey', 'grey', 'red'),
                 colAlpha        = 1,
                 selectLab       = c('PAGE4', 'SMAGP', 'BCAM', 'DUSP9', 'SPINT1', 'EGR1', 'CLU', 'IFI27', 'ISG15', 'EPCAM', 'HLA-G'),
                 xlim            = c(-4.1, 4.1),
                 gridlines.major = FALSE,
                 drawConnectors  = TRUE, 
                 gridlines.minor = FALSE) + NoLegend () + ggtitle (NULL)
```

## SCENIC analyses (run 5 times each)

### Maternal-Placental Interface (in vivo) CTB (repeat 5 times)

```{r}
CTBs    <- subset (Trophoblasts, idents = c("CTB 1", "CTB 2", "CTB 3", "CTB 4"))
MPI     <- subset (CTBs, subset = Model == "in vivo")
MPI.CTB <- subset (invivo,  downsample = 500)
MPI.CTB

table (Idents (MPI.CTB))
rm    (Trophoblasts)
rm    (MPI)
```

```{r}
DefaultAssay (MPI.CTB) <- 'RNA'

exprMat  <- MPI.CTB@assays$RNA@data
cellInfo <- MPI.CTB@meta.data

dim (exprMat)
```

```{r}
org            <- "hgnc"
dbDir          <- "/path_to_MPI_directory"
myDatasetTitle <- "Trophoblast_SCENIC"

data (defaultDbNames)

           dbs <- defaultDbNames[[org]]
 scenicOptions <- initializeScenic (org          = org,
                                    dbDir        = dbDir,
                                    dbs          = dbs,
                                    datasetTitle = myDatasetTitle,
                                    nCores       = 32)
 
scenicOptions@settings$verbose <- TRUE
scenicOptions@settings$nCores  <- 32
```

```{r, Edit code for geneFiltering}
geneFiltering <- function (exprMat, scenicOptions,
                           minCountsPerGene = 3*.01*ncol (exprMat),
                           minSamples       = ncol (exprMat)*.01)
{
  # Load options: outFile_genesKept and dbFilePath
  outFile_genesKept <- NULL
  dbFilePath        <- NULL
  if (class (scenicOptions) == "ScenicOptions")
  {
    dbFilePath        <- getDatabases (scenicOptions)[[1]]
    outFile_genesKept <- getIntName   (scenicOptions, "genesKept")
  }else{
    dbFilePath        <- scenicOptions[["dbFilePath"]]
    outFile_genesKept <- scenicOptions[["outFile_genesKept"]]
  }
  if (is.null (dbFilePath)) stop ("dbFilePath")
  
  # Check expression matrix (e.g. not factor)
  if (is.data.frame (exprMat)) 
  {
    supportedClasses <- paste (gsub ("AUCell_buildRankings,", "", methods ("AUCell_buildRankings")), collapse = ", ")
    supportedClasses <- gsub  ("-method", "", supportedClasses)
    
    stop ("'exprMat' should be one of the following classes: ", supportedClasses, 
          "(data.frames are not supported. Please, convert the expression matrix to one of these classes.)")
  }
  if (any (table (rownames (exprMat)) > 1))
    stop ("The rownames (gene id/name) in the expression matrix should be unique.")
  
  # Calculate stats
  nCountsPerGene <- rowSums (exprMat,     na.rm = TRUE)
  nCellsPerGene  <- rowSums (exprMat > 0, na.rm = TRUE)
  
  ## Show info
  message ("Maximum value in the expression matrix: ", max    (exprMat, na.rm = TRUE))
  message ("Ratio of detected vs non-detected: ",      signif (sum (exprMat > 0, na.rm = TRUE) / sum (exprMat == 0, na.rm = TRUE), 2))
  message ("Number of counts (in the dataset units) per gene:")
  print   (summary (nCountsPerGene))
  message ("Number of cells in which each gene is detected:")
  print   (summary (nCellsPerGene))

  ## Filter
  message ("\nNumber of genes left after applying the following filters (sequential):")
  
  # First filter
  # minCountsPerGene <- 3*.01*ncol (exprMat)
  genesLeft_minReads <- names (nCountsPerGene)[which (nCountsPerGene > minCountsPerGene)]
  message ("\t", length (genesLeft_minReads), "\tgenes with counts per gene > ", minCountsPerGene)
  
  # Second filter
  # minSamples       <- ncol (exprMat)*.01
  nCellsPerGene2     <- nCellsPerGene[genesLeft_minReads]
  genesLeft_minCells <- names (nCellsPerGene2)[which(nCellsPerGene2 > minSamples)]
  message ("\t", length (genesLeft_minCells), "\tgenes detected in more than ", minSamples," cells")
  
  # Exclude genes missing from database:
  library (RcisTarget)
  motifRankings                  <- importRankings (dbFilePath) # either one, they should have the same genes
  genesInDatabase                <- colnames       (getRanking (motifRankings))
  genesLeft_minCells_inDatabases <- genesLeft_minCells[which (genesLeft_minCells %in% genesInDatabase)]
  message ("\t", length (genesLeft_minCells_inDatabases), "\tgenes available in RcisTarget database")
  
  genesKept <- genesLeft_minCells_inDatabases
  if (!is.null (outFile_genesKept)){ 
      saveRDS  (genesKept, file=outFile_genesKept)
    if (getSettings (scenicOptions, "verbose")) message ("Gene list saved in ", outFile_genesKept)
  }
  return (genesKept)
}
```

```{r}
genesKept <- geneFiltering (exprMat,
                            scenicOptions    = scenicOptions,
                            minCountsPerGene = 3*.01*ncol (exprMat),
                            minSamples       = ncol       (exprMat)*.01)
```

```{r}
exprMat_filtered <- exprMat[genesKept, ]
dim (exprMat_filtered)
rm  (exprMat)
```

```{r, Convert to matrix for runCorrelation}
exprMat_filtered <- as.matrix (exprMat_filtered)
```

```{r}
runCorrelation (exprMat_filtered, scenicOptions)
runGenie3      (exprMat_filtered, scenicOptions)

scenicOptions <- runSCENIC_1_coexNetwork2modules (scenicOptions)
scenicOptions <- runSCENIC_2_createRegulons      (scenicOptions)
scenicOptions <- runSCENIC_3_scoreCells          (scenicOptions, exprMat_filtered)
scenicOptions <- runSCENIC_4_aucell_binarize     (scenicOptions)
regulons      <- loadInt                         (scenicOptions, "aucell_regulonAUC")
```

```{r}
regulonAUC                        <- loadInt  (scenicOptions, "aucell_regulonAUC")
regulonAUC                        <- regulonAUC[onlyNonDuplicatedExtended (rownames (regulonAUC)), ]
regulonActivity_byCellType        <- sapply   (split (rownames (cellInfo), cellInfo$seurat_clusters),
                                     function (cells) rowMeans (getAUC (regulonAUC)[ , cells]))
regulonActivity_byCellType_Scaled <- t        (scale (t (regulonActivity_byCellType), center = TRUE, scale = TRUE))
regulonActivity_byCellType_Scaled <- regulonActivity_byCellType_Scaled[ , -c(3,5,6,7,8,9,11)]
```

```{r, Supplemental figure 3, only one representative heatmap is shown in the paper}
ComplexHeatmap::Heatmap (regulonActivity_byCellType_Scaled, name = "Regulon activity")
```

```{r}
minPerc <- 0.5
binaryRegulonActivity                <- loadInt (scenicOptions, "aucell_binary_nonDupl")
cellInfo_binarizedCells              <- cellInfo[which (rownames (cellInfo)%in% colnames (binaryRegulonActivity)), , drop = FALSE]
regulonActivity_byCellType_Binarized <- sapply  (split (rownames (cellInfo_binarizedCells), cellInfo_binarizedCells$seurat_clusters), 
                                                 function (cells) rowMeans (binaryRegulonActivity[ , cells, drop = FALSE]))
regulonActivity_byCellType_Binarized <- regulonActivity_byCellType_Binarized[ , -c(3,5,6,7,8,9,11)]
binaryActPerc_subset                 <- regulonActivity_byCellType_Binarized[which (rowSums (regulonActivity_byCellType_Binarized>minPerc) > 0), ]
```

```{r, Supplemental figure 3, only one representative heatmap is shown in the paper}
ComplexHeatmap::Heatmap (binaryActPerc_subset, name="Regulon activity (%)", col = c("white","pink","red"))
```

```{r}
regulonAUC                        <- regulonAUC[onlyNonDuplicatedExtended (rownames (regulonAUC)), ]
regulonActivity_byCellType_Scaled <- t (scale (t (regulonActivity_byCellType), center = TRUE, scale = TRUE))
regulonActivity_byCellType_Scaled <- regulonActivity_byCellType_Scaled[ , -c(3,5,6,7,8,9,11)]
regulonActivity_byCellType_Scaled
```

```{r, Table S7A}
write.xlsx (regulonActivity_byCellType_Scaled, file = "Ranked_MPI_CTB_regulons_run#.xlsx")
```

### TBpOrg (in vitro) CTB (repeat 5 times)

```{r}
TBpOrg.CTB <- subset (CTBs, subset = Model == "TBpOrg")

table (Idents (TBpOrg.CTB))
```

```{r}
DefaultAssay (TBpOrg.CTB) <- 'RNA'

exprMat  <- TBpOrg.CTB@assays$RNA@data
cellInfo <- TBpOrg.CTB@meta.data

dim (exprMat)
```

```{r}
org            <- "hgnc"
dbDir          <- "/path_to_TBpOrg_directory"
myDatasetTitle <- "Trophoblast_SCENIC"

data (defaultDbNames)

           dbs <- defaultDbNames[[org]]
 scenicOptions <- initializeScenic (org          = org,
                                    dbDir        = dbDir,
                                    dbs          = dbs,
                                    datasetTitle = myDatasetTitle,
                                    nCores       = 32) 

scenicOptions@settings$verbose <- TRUE
scenicOptions@settings$nCores  <- 32
```

```{r}
genesKept <- geneFiltering (exprMat,
                            scenicOptions    = scenicOptions,
                            minCountsPerGene = 3*.01*ncol (exprMat),
                            minSamples       = ncol       (exprMat)*.01)
```

```{r}
exprMat_filtered <- exprMat[genesKept, ]
dim (exprMat_filtered)
rm  (exprMat)
```

```{r, Convert to matrix for runCorrelation}
exprMat_filtered <- as.matrix (exprMat_filtered)
```

```{r}
runCorrelation (exprMat_filtered, scenicOptions)
runGenie3      (exprMat_filtered, scenicOptions)

scenicOptions <- runSCENIC_1_coexNetwork2modules (scenicOptions)
scenicOptions <- runSCENIC_2_createRegulons      (scenicOptions)
scenicOptions <- runSCENIC_3_scoreCells          (scenicOptions, exprMat_filtered)
scenicOptions <- runSCENIC_4_aucell_binarize     (scenicOptions)
regulons      <- loadInt                         (scenicOptions, "aucell_regulonAUC")
```

```{r}
regulonAUC                        <- loadInt  (scenicOptions, "aucell_regulonAUC")
regulonAUC                        <- regulonAUC[onlyNonDuplicatedExtended (rownames (regulonAUC)), ]
regulonActivity_byCellType        <- sapply   (split (rownames (cellInfo), cellInfo$seurat_clusters),
                                     function (cells) rowMeans (getAUC (regulonAUC)[ , cells]))
regulonActivity_byCellType_Scaled <- t        (scale (t (regulonActivity_byCellType), center = TRUE, scale = TRUE))
regulonActivity_byCellType_Scaled <- regulonActivity_byCellType_Scaled[ , -c(3,5,6,7,8,9,11)]
```

```{r, Supplemental figure 3, only one representative heatmap is shown in the paper}
ComplexHeatmap::Heatmap (regulonActivity_byCellType_Scaled, name = "Regulon activity")
```

```{r}
minPerc <- 0.5
binaryRegulonActivity                <- loadInt (scenicOptions, "aucell_binary_nonDupl")
cellInfo_binarizedCells              <- cellInfo[which (rownames (cellInfo) %in% colnames (binaryRegulonActivity)), , drop = FALSE]
regulonActivity_byCellType_Binarized <- sapply  (split (rownames (cellInfo_binarizedCells), cellInfo_binarizedCells$seurat_clusters), 
                                                 function (cells) rowMeans(binaryRegulonActivity[ , cells, drop = FALSE]))
regulonActivity_byCellType_Binarized <- regulonActivity_byCellType_Binarized[, -c(3,5,6,7,8,9,11)]
binaryActPerc_subset                 <- regulonActivity_byCellType_Binarized[which (rowSums (regulonActivity_byCellType_Binarized > minPerc) > 0), ]
```

```{r, Supplemental figure 3, only one representative heatmap is shown in the paper}
ComplexHeatmap::Heatmap (binaryActPerc_subset, name="Regulon activity (%)", col = c("white","pink","red"))
```

```{r}
regulonAUC                        <- regulonAUC[onlyNonDuplicatedExtended (rownames (regulonAUC)), ]
regulonActivity_byCellType_Scaled <- t (scale (t (regulonActivity_byCellType), center = TRUE, scale = TRUE))
regulonActivity_byCellType_Scaled <- regulonActivity_byCellType_Scaled[ , -c(3,5,6,7,8,9,11)]
regulonActivity_byCellType_Scaled
```

```{r, Table S7B}
write.xlsx (regulonActivity_byCellType_Scaled, file = "Ranked_TBpOrg_CTB_regulons_run#.xlsx")
```

### TBsOrg (in vitro) CTB (repeat 5 times)

```{r}
TBsOrg.CTB <- subset (CTBs, subset = Model == "TBsOrg")

table (Idents (TBsOrg.CTB))
```

```{r}
DefaultAssay (TBsOrg.CTB) <- 'RNA'

exprMat  <- TBsOrg.CTB@assays$RNA@data
cellInfo <- TBsOrg.CTB@meta.data

dim (exprMat)
```

```{r}
org            <- "hgnc"
dbDir          <- "/path_to_TBsOrg_directory"
myDatasetTitle <- "Trophoblast_SCENIC"

data (defaultDbNames)

           dbs <- defaultDbNames[[org]]
 scenicOptions <- initializeScenic (org          = org,
                                    dbDir        = dbDir,
                                    dbs          = dbs,
                                    datasetTitle = myDatasetTitle,
                                    nCores       = 32) 

scenicOptions@settings$verbose <- TRUE
scenicOptions@settings$nCores  <- 32
```

```{r}
genesKept <- geneFiltering (exprMat,
                            scenicOptions    = scenicOptions,
                            minCountsPerGene = 3*.01*ncol (exprMat),
                            minSamples       = ncol       (exprMat)*.01)
```

```{r}
exprMat_filtered <- exprMat[genesKept, ]
dim (exprMat_filtered)
rm  (exprMat)
```

```{r, Convert to matrix for runCorrelation}
exprMat_filtered <- as.matrix (exprMat_filtered)
```

```{r}
runCorrelation (exprMat_filtered, scenicOptions)
runGenie3      (exprMat_filtered, scenicOptions)

scenicOptions <- runSCENIC_1_coexNetwork2modules (scenicOptions)
scenicOptions <- runSCENIC_2_createRegulons      (scenicOptions)
scenicOptions <- runSCENIC_3_scoreCells          (scenicOptions, exprMat_filtered)
scenicOptions <- runSCENIC_4_aucell_binarize     (scenicOptions)
regulons      <- loadInt                         (scenicOptions, "aucell_regulonAUC")
```

```{r}
regulonAUC                        <- loadInt  (scenicOptions, "aucell_regulonAUC")
regulonAUC                        <- regulonAUC[onlyNonDuplicatedExtended (rownames (regulonAUC)), ]
regulonActivity_byCellType        <- sapply   (split (rownames (cellInfo), cellInfo$seurat_clusters),
                                     function (cells) rowMeans (getAUC (regulonAUC)[ , cells]))
regulonActivity_byCellType_Scaled <- t        (scale (t (regulonActivity_byCellType), center = TRUE, scale = TRUE))
regulonActivity_byCellType_Scaled <- regulonActivity_byCellType_Scaled[ , -c(3,5,6,7,8,9,11)]
```

```{r, Supplemental figure 3, only one representative heatmap is shown in the paper}
ComplexHeatmap::Heatmap (regulonActivity_byCellType_Scaled, name = "Regulon activity")
```

```{r}
minPerc <- 0.5
binaryRegulonActivity                <- loadInt (scenicOptions, "aucell_binary_nonDupl")
cellInfo_binarizedCells              <- cellInfo[which (rownames (cellInfo) %in% colnames (binaryRegulonActivity)), , drop = FALSE]
regulonActivity_byCellType_Binarized <- sapply  (split (rownames (cellInfo_binarizedCells), cellInfo_binarizedCells$seurat_clusters), 
                                                 function (cells) rowMeans (binaryRegulonActivity[ , cells, drop = FALSE]))
regulonActivity_byCellType_Binarized <- regulonActivity_byCellType_Binarized[ , -c(3,5,6,7,8,9,11)]
binaryActPerc_subset                 <- regulonActivity_byCellType_Binarized[which (rowSums (regulonActivity_byCellType_Binarized > minPerc) > 0), ]
```

```{r, Supplemental figure 3, only one representative heatmap is shown in the paper}
ComplexHeatmap::Heatmap (binaryActPerc_subset, name="Regulon activity (%)", col = c("white","pink","red"))
```

```{r}
regulonAUC                        <- regulonAUC[onlyNonDuplicatedExtended (rownames (regulonAUC)), ]
regulonActivity_byCellType_Scaled <- t (scale (t (regulonActivity_byCellType), center = TRUE, scale = TRUE))
regulonActivity_byCellType_Scaled <- regulonActivity_byCellType_Scaled[ , -c(3,5,6,7,8,9,11)]
regulonActivity_byCellType_Scaled
```

```{r, Table S7C}
write.xlsx (regulonActivity_byCellType_Scaled, file = "Ranked_TBsOrg_CTB_regulons_run#.xlsx")
```

### Integrated (all models) trophoblasts (repeat 5 times)

```{r}
Troph.sub  <- subset (Trophoblasts, downsample = 500)

table (Idents (Troph.sub))
rm    (Trophoblasts)
```

```{r}
DefaultAssay (Troph.sub) <- 'RNA'

exprMat  <- Troph.sub@assays$RNA@data
cellInfo <- Troph.sub@meta.data

dim (exprMat)
```

```{r}
org            <- "hgnc"
dbDir          <- "/path_to_Trophoblast_directory"
myDatasetTitle <- "Trophoblast_SCENIC"

data (defaultDbNames)

           dbs <- defaultDbNames[[org]]
 scenicOptions <- initializeScenic (org          = org,
                                    dbDir        = dbDir,
                                    dbs          = dbs,
                                    datasetTitle = myDatasetTitle,
                                    nCores       = 32) 

scenicOptions@settings$verbose <- TRUE
scenicOptions@settings$nCores  <- 32
```

```{r}
genesKept <- geneFiltering (exprMat,
                            scenicOptions    = scenicOptions,
                            minCountsPerGene = 3*.01*ncol (exprMat),
                            minSamples       = ncol       (exprMat)*.01)
```

```{r}
exprMat_filtered <- exprMat[genesKept, ]
dim (exprMat_filtered)
rm  (exprMat)
```

```{r, Convert to matrix for runCorrelation}
exprMat_filtered <- as.matrix (exprMat_filtered)
```

```{r}
runCorrelation (exprMat_filtered, scenicOptions)
runGenie3      (exprMat_filtered, scenicOptions)

scenicOptions <- runSCENIC_1_coexNetwork2modules (scenicOptions)
scenicOptions <- runSCENIC_2_createRegulons      (scenicOptions)
scenicOptions <- runSCENIC_3_scoreCells          (scenicOptions, exprMat_filtered)
scenicOptions <- runSCENIC_4_aucell_binarize     (scenicOptions)
regulons      <- loadInt                         (scenicOptions, "aucell_regulonAUC")
```

```{r}
regulonAUC                        <- regulonAUC[onlyNonDuplicatedExtended (rownames (regulonAUC)), ]
regulonActivity_byCellType        <- sapply   (split (rownames (cellInfo), cellInfo$seurat_clusters),
                                     function (cells) rowMeans (getAUC (regulonAUC)[ , cells]))
regulonActivity_byCellType_Scaled <- t        (scale (t (regulonActivity_byCellType), center = TRUE, scale = TRUE))
```

```{r, Supplemental figure 3, only one representative heatmap is shown in the paper}
ComplexHeatmap::Heatmap (regulonActivity_byCellType_Scaled, name = "Regulon activity")
```

```{r}
minPerc <- 0.5
binaryRegulonActivity                <- loadInt (scenicOptions, "aucell_binary_nonDupl")
cellInfo_binarizedCells              <- cellInfo[which (rownames (cellInfo) %in% colnames (binaryRegulonActivity)), , drop = FALSE]
regulonActivity_byCellType_Binarized <- sapply  (split (rownames (cellInfo_binarizedCells), cellInfo_binarizedCells$seurat_clusters), 
                                                 function (cells) rowMeans (binaryRegulonActivity[ , cells, drop = FALSE]))
binaryActPerc_subset                 <- regulonActivity_byCellType_Binarized[which (rowSums (regulonActivity_byCellType_Binarized > minPerc) > 0), ]
```

```{r, Supplemental figure 3, only one representative heatmap is shown in the paper}
ComplexHeatmap::Heatmap (binaryActPerc_subset, name = "Regulon activity (%)", col = c("white","pink","red"))
```

```{r}
regulonAUC                        <- regulonAUC[onlyNonDuplicatedExtended (rownames (regulonAUC)), ]
regulonActivity_byCellType_Scaled <- t (scale (t (regulonActivity_byCellType), center = T, scale = T))
regulonActivity_byCellType_Scaled
```

```{r, Table S7D}
write.xlsx (regulonActivity_byCellType_Scaled, file = "Ranked_Integrated.Trophoblasts_regulons_run#.xlsx")
```

## Plotting the SCENIC results

```{r, Figure 3C and 3D regulon transcription factor dot plot}
         Idents (CTBs) <- CTBs$Model
table   (Idents (CTBs))

DotPlot (CTBs,
         features  = c("SRF",
                       "HSF1",
                       "CEBPD",
                       "ESRRA",
                       "CEBPB",
                       "GATA2",
                       "GATA3",
                       "TFAP2A",
                       "TFAP2C",
                       "ELF5",
                       "TP63",
                       "SMARCA4",
                       "ATF1",
                       "TBL1XR1",
                       "FOSL2",
                       "NFIL3",
                       "ETV5",
                       "ATF4",
                       "EZH2",
                       "XBP1",
                       "DDIT3",
                       "EGR1",
                       "E2F4",
                       "FOXP1",
                       "FOXO4",
                       "IRF6",
                       "TFAP2B",
                       "PBX1",
                       "TEAD4"),
         assay     = "RNA",
         cols      = c("lightgrey", "red"),
         col.min   = 0,
         col.max   = 1,
         dot.min   = 0,
         dot.scale = 15) +
theme   (axis.text.x = element_text(angle = 90))
```

```{r, Figure 3D FOXO4 UMAP}
FeaturePlot (Trophoblasts, features = c("FOXO4"), label = FALSE, pt.size = 3, min.cutoff = 0, order = TRUE, cols = c("grey", "red")) + NoLegend ()
```

# TBsOrg CTB Development

```{r, Figure 3F subset}
CTB_TSC <- subset (TBsOrg, idents = c("CTB", "TSC"))
CTB_TSC
```

```{r, Figure 3F UMAP}
UMAPPlot (CTB_TSC, label = TRUE, pt.size = 3) + NoLegend ()
```

```{r, Figure 3F Pseudotime analysis}
CTB_TSC.monocle3 <- as.cell_data_set      (CTB_TSC, assay = "RNA")
CTB_TSC.monocle3 <- estimate_size_factors (CTB_TSC.monocle3)
CTB_TSC.monocle3 <- cluster_cells         (CTB_TSC.monocle3,
                                           reduction_method = c("UMAP"),
                                           cluster_method   = "leiden",
                                           partition_qval   = 0.05,
                                           resolution       = 0.0001)
p1  <- plot_cells (CTB_TSC.monocle3,                               show_trajectory_graph = FALSE)
p2  <- plot_cells (CTB_TSC.monocle3, color_cells_by = "partition", show_trajectory_graph = FALSE)
wrap_plots        (p1, p2)
```

```{r, Figure 3F Pseudotime analysis}
CTB_TSC.monocle3 <- learn_graph (CTB_TSC.monocle3, use_partition = FALSE, close_loop = TRUE)
```

```{r, Figure 3F Pseudotime UMAP (left)}
plot_cells (CTB_TSC.monocle3,
            color_cells_by                = "celltype",
            label_cell_groups             = FALSE,
            label_leaves                  = FALSE,
            label_branch_points           = FALSE,
            graph_label_size              = 1.5,
            cell_size                     = 3,
            trajectory_graph_color        = "black",
            trajectory_graph_segment_size = 1.25) + NoLegend ()
```

```{r, Figure 3F Pseudotime UMAP (right)}
CTB_TSC.monocle3 <- order_cells (CTB_TSC.monocle3)

plot_cells (CTB_TSC.monocle3,
            color_cells_by                = "pseudotime",
            label_cell_groups             = FALSE,
            label_leaves                  = FALSE,
            label_branch_points           = FALSE,
            graph_label_size              = 1.5,
            cell_size                     = 3,
            trajectory_graph_color        = "black",
            trajectory_graph_segment_size = 1.25) + NoLegend ()
```

```{r, Figure 3G, rename cell idents}
MPI <- RenameIdents (MPI,
                     'CTB'  = "CTB_TB",
                     'cCTB' = "cCTB_TB",
                     'EVT'  = "EVT_TB",
                     'SCTp' = "SCTp_TB")
MPI$celltype <- Idents (MPI)

TBpOrg <- RenameIdents (TBpOrg,
                        'CTB'  = "CTB_TP",
                        'cCTB' = "cCTB_TP",
                        'EVT'  = "EVT_TP",
                        'SCTp' = "SCTp_TP",
                        'EVT'  = "TSC_TP")
TBpOrg$celltype <- Idents (TBpOrg)

TBsOrg <- RenameIdents (TBsOrg,
                        'TSC'  = "TSC_TS",
                        'cCTB' = "cCTB_TS",
                        'SCTp' = "SCTp_TS",
                        'CTB'  = "CTB_TS",
                        'EVT'  = "EVT_TS" )
TBsOrg$celltype <- Idents (TBsOrg)
```

```{r, Figure 3G, subset}
CTB_TB <- subset (MPI, idents = c("CTB_TB"))
CTB_TB

CTB_TP <- subset (TBpOrg, idents = c("CTB_TP"))
CTB_TP

CTB_TSC_TS <- subset (TBsOrg, idents = c("CTB_TS", "TSC_TS"))
CTB_TSC_TS
```

```{r, Figure 3G, dataset preparation}
CTB_TSC_COMP <- merge (CTB_TB,   y = c(CTB_TP, CTB_TSC_TS),
                       add.cell.ids = c("TB", "TP", "TS"),
                       project      = "PCC",
                       merge.data   = TRUE)
CTB_TSC_COMP
```

```{r, Figure 3G, heatmap}
av.exp                <- AverageExpression (CTB_TSC_COMP)$RNA
cor.exp               <- as.data.frame     (cor (av.exp))
cor.exp$x             <- rownames          (cor.exp)
cor.df                <- tidyr::gather     (data = cor.exp, y, correlation, c('CTB_TB', 'CTB_TP', 'CTB_TS', 'TSC_TS'))
cor.df

my_levels <- c("CTB_TB", "CTB_TP", "CTB_TS", "TSC_TS")
cor.df$x  <- factor (cor.df$x, levels = my_levels)
cor.df$y  <- factor (cor.df$y, levels = my_levels)

ggplot               (cor.df, aes (x, y, fill = correlation)) +
geom_tile            () +
scale_fill_gradientn (limits = c(0, 1), colours = colorRampPalette (rev (brewer.pal (n = 7, name = "RdYlBu"))) (100))
```

```{r, Figure 3H, differential gene expression analysis}
DefaultAssay (CTB_TSC) <- "RNA"

CTB.TSC.response <- FindMarkers (CTB_TSC,
                                 ident.1         = c("CTB"),
                                 ident.2         = c("TSC"),
                                 test.use        = "MAST",
                                 logfc.threshold = -Inf,
                                 min.pct         = -Inf,
                                 min.diff.pct    = -Inf,
                                 verbose         = FALSE)
                     head       (CTB.TSC.response, n = 100)
```

```{r, Table S6A}
library ("xlsx")
write.xlsx (CTB.TSC.response, file = "TBsOrg_CTB_TSC_DGE.xlsx")
```

```{r, Figure 3H, volcano plot}
EnhancedVolcano (CTB.TSC.response,
                 lab             = rownames (CTB.TSC.response),
                 x               = "avg_log2FC",
                 y               = "p_val_adj",
                 FCcutoff        = 1.00,
                 pCutoff         = 10e-50,
                 pointSize       = 4.0,
                 col             = c('grey', 'grey', 'grey', 'red'),
                 colAlpha        = 1,
                 selectLab       = c('PAGE4', 'SERPINF1', 'PEG10', 'VGLL1', 'PHLDA2', 'CD9', 'FTL', 'TMSB4X', 'PSG5', 'NEAT1', 'ITGA2'),
                 xlim            = c(-4.1, 4.1),
                 ylim            = c(0, 325),
                 gridlines.major = FALSE,
                 drawConnectors  = TRUE, 
                 gridlines.minor = FALSE) + NoLegend () + ggtitle (NULL)
```
