---
title:  "MODEL_COMPARISON_FIGURE4"
author: "Matthew J Shannon"
date:   "18/01/2024"
output: html_document
---

# Load in all dependencies

```{r, include = FALSE}
devtools::install_github ('chris-mcginnis-ucsf/DoubletFinder')
devtools::install_github ('satijalab/seurat-wrappers')
remotes::install_github  ("mojaveazure/seurat-disk")
devtools::install_github ("velocyto-team/velocyto.R")
BiocManager::install     ("pcaMethods")
BiocManager::install     ('MAST')
BiocManager::install     ('org.Hs.eg.db')
BiocManager::install     (version = "3.12")
BiocManager::install     ('Seurat')
BiocManager::install     ('readxl')
BiocManager::install     ('modes')
BiocManager::install     ('pheatmap')
BiocManager::install     ('limma')
BiocManager::install     ('clustree')
BiocManager::install     ('clusterProfiler')
BiocManager::install     ('EnhancedVolcano')
install.packages         ('fields')
install.packages         ('plotly')
install.packages         ("VennDiagram")
library                  (Seurat)
library                  (readr)
library                  (tidyverse)
library                  (Matrix)
library                  (dplyr)
library                  (ggplot2)
library                  (ggpubr)
library                  (cowplot)
library                  (tibble)
library                  (readxl)
library                  (sctransform)
library                  (fields)
library                  (KernSmooth)
library                  (ROCR)
library                  (parallel)
library                  (reshape2)
library                  (pheatmap)
library                  (DoubletFinder)
library                  (limma)
library                  (SeuratWrappers)
library                  (SeuratDisk)
library                  (plotly)
library                  (clustree)
library                  (velocyto.R)
library                  (MAST)
library                  (EnhancedVolcano)
library                  (clusterProfiler)
library                  (AnnotationDbi)
library                  (org.Hs.eg.db)
library                  (VennDiagram)
library                  (RColorBrewer)
library                  (EnhancedVolcano)
library                  (xlsx)
gc                       ()
sessionInfo              ()
```

# Load the integrated and model-specific trophoblast Seurat objects

```{r}
# Integrated (all models) trophoblast object
load         (file = "~/Desktop/FINAL_Trophoblast_Data.Rdata")
DefaultAssay (Trophoblasts) <- "integrated"

# MPI (in vivo) trophoblast object
load         (file = "~/Desktop/MPI.Rdata")
DefaultAssay (MPI) <- "integrated"

# TBpOrg (in vitro) trophoblast object
load         (file = "~/Desktop/TBpOrg.Rdata")
DefaultAssay (TBpOrg) <- "integrated"

# TBsOrg (in vitro) trophoblast object
load         (file = "~/Desktop/TBsOrg.Rdata")
DefaultAssay (TBsOrg) <- "integrated"
```

# Figure 6

## Villous Differentiation Branch Comparison

```{r, Figure 4A, Integrated (in vivo + in vitro) villous differentiation branch subset}
Villous <- subset (Trophoblasts, idents = c("CTB 1", "CTB 2", "CTB 3", "CTB 4", "SCTp 1", "SCTp 2"))
```

```{r, Figure 4A, Integrated (in vivo + in vitro) villous differentiation branch UMAP}
UMAPPlot (Villous, label = TRUE, pt.size = 3) + NoLegend ()
```

```{r, Figure 4B subset}
SCTp <- subset (Trophoblasts, idents = c("SCTp 1", "SCTp 2"))
```

```{r, Figure 4B heatmap}
Idents (SCTp) <- SCTp@meta.data$Model
av.exp        <- AverageExpression (SCTp)$RNA
cor.exp       <- as.data.frame     (cor (av.exp))
cor.exp$x     <- rownames          (cor.exp)
cor.df        <- tidyr::gather     (data = cor.exp, y, correlation, c('MPI', 'TBpOrg', 'TBsOrg'))
cor.df

my_levels <- c("MPI", "TBpOrg", "TBsOrg")
cor.df$x  <- factor (cor.df$x, levels = my_levels)
cor.df$y  <- factor (cor.df$y, levels = my_levels)

ggplot               (cor.df, aes (x, y, fill = correlation)) +
geom_tile            () +
scale_fill_gradientn (limits = c(0, 1), colours = colorRampPalette (rev (brewer.pal (n = 7, name = "RdYlBu"))) (100))
```

## SCTp Differential gene expression analysis

```{r, Figure 4E MPI vs TBpOrg differential gene expression analysis}
DefaultAssay (SCTp) <- "RNA"

Idents (SCTp) <- SCTp@meta.data$Model

TBpOrg.SCTp.response <- FindMarkers (SCTp,
                                  ident.1         = c("MPI"),
                                  ident.2         = c("TBpOrg"),
                                  test.use        = "MAST",
                                  logfc.threshold = -Inf,
                                  min.pct         = -Inf,
                                  min.diff.pct    = -Inf,
                                  verbose         = FALSE)
                     head        (TBpOrg.SCTp.response, n = 100)
```

```{r}
write.xlsx (TBpOrg.SCTp.response, file = "MPI_TBpOrg_SCTp_DGE.xlsx")
```

```{r, Figure 4E MPI vs TBpOrg volcano plot}
EnhancedVolcano (TBpOrg.SCTp.response,
                 lab             = rownames (TBpOrg.SCTp.response),
                 x               = "avg_log2FC",
                 y               = "p_val_adj",
                 FCcutoff        = 1.00,
                 pCutoff         = 10e-50,
                 pointSize       = 4.0,
                 col             = c('grey', 'grey', 'grey', 'red'),
                 colAlpha        = 1,
                 selectLab       = c('FOS', 'SLC22A11', 'BSG', 'GPX3', 'FXYD3', 'ARG2', 'BRI3', 'SLC7A8', 'TIMP2', 'SCD', 'ISG15', 'IFI27'),
                 xlim            = c(-6, 6),
                 ylim            = c(0, 325),
                 gridlines.major = FALSE,
                 drawConnectors  = TRUE, 
                 gridlines.minor = FALSE) + NoLegend () + ggtitle (NULL)
```

```{r, Figure 4E MPI vs TBsOrg differential gene expression analysis}
DefaultAssay (SCTp) <- "RNA"

Idents (SCTp) <- SCTp@meta.data$Model

TBsOrg.SCTp.response <- FindMarkers (SCTp,
                                   ident.1         = c("MPI"),
                                   ident.2         = c("TBsOrg"),
                                   test.use        = "MAST",
                                   logfc.threshold = -Inf,
                                   min.pct         = -Inf,
                                   min.diff.pct    = -Inf,
                                   verbose         = FALSE)
                      head        (TBsOrg.SCTp.response, n = 100)
```

```{r}
write.xlsx (TBsOrg.SCTp.response, file = "MPI_TBsOrg_SCTp_DGE.xlsx")
```

```{r, Figure 4E MPI vs TBsOrg volcano plot}
EnhancedVolcano (TBsOrg.SCTp.response,
                 lab             = rownames (TBsOrg.SCTp.response),
                 x               = "avg_log2FC",
                 y               = "p_val_adj",
                 FCcutoff        = 1.00,
                 pCutoff         = 10e-50,
                 pointSize       = 4.0,
                 col             = c('grey', 'grey', 'grey', 'red'),
                 colAlpha        = 1,
                 selectLab       = c('PAGE4', 'FOS', 'VGLL1', 'SLC22A11', 'SLC26A2', 'DUSP9', 'GPX3', 'APOE', 'TIMP2', 'FTL', 'ISG15', 'NEAT1', 'IFI27', 'SCD', 'CGB'),
                 xlim            = c(-6, 6),
                 ylim            = c(0, 325),
                 gridlines.major = FALSE,
                 drawConnectors  = TRUE, 
                 gridlines.minor = FALSE) + NoLegend () + ggtitle (NULL)
```

## Model-specific villous differentiation branch regulon transcription factor dot plots

```{r, Figure 4F, dot plot (top)}
# Comparing Maternal-Placental Interface scVelo-informed gene driver expression across each trophoblast model
DotPlot (SCTp,
         features = c("HES1",
                      "PITX1",
                      "MXD1",
                      "ETV4",
                      "TFAP2A",
                      "TFAP2C",
                      "NFKB2",
                      "ZNF165",
                      "ZNF143",
                      "MAFF",
                      "MAFK",
                      "MAFG",
                      "GMEB1",
                      "CREM",
                      "CEBPA",
                      "NR2F6"),
         assay     = "RNA",
         cols      = c("lightgrey", "red"),
         col.min   = 0,
         col.max   = 1,
         dot.min   = 0,
         dot.scale = 15)
```

## State-specific villous differentiation branch regulon transcription factor dot plots

```{r, Figure 4F, dot plot (bottom)}
# Plotting transcription factors identified in Table S7 across each SCTp state
Idents  (SCTp) <- SCTp$Seurat_clusters
DotPlot (SCTp,
         features = c("HES1",
                      "PITX1",
                      "MXD1",
                      "ETV4",
                      "TFAP2A",
                      "TFAP2C",
                      "NFKB2",
                      "ZNF165",
                      "ZNF143",
                      "MAFF",
                      "MAFK",
                      "MAFG",
                      "GMEB1",
                      "CREM",
                      "CEBPA",
                      "NR2F6"),
         assay     = "RNA",
         cols      = c("lightgrey", "red"),
         col.min   = 0,
         col.max   = 1,
         dot.min   = 0,
         dot.scale = 15)
```

# Supplemental Figure 6

### SCTp gene expression comparison

```{r, Supplemental figure 4A renaming cluster idents}
Trophoblasts <- RenameIdents (Trophoblasts,
                              'CTB 1'  = "CTB",
                              'CTB 2'  = "CTB",
                              'CTB 3'  = "CTB",
                              'CTB 4'  = "CTB",
                              'TSC'    = "TSC",
                              'cCTB 1' = "cCTB",
                              'cCTB 2' = "cCTB",
                              'cCTB 3' = "cCTB",
                              'EVT'    = "EVT",
                              'SCTp 1' = "SCTp 1",
                              'SCTp 2' = "SCTp 2")

Trophoblasts$celltype <- Idents (Trophoblasts)
```

```{r, Supplemental figure 4A}
Trophoblasts$celltype.model  <- paste  (Idents (Trophoblasts), Trophoblasts$Model, sep = "_")
Trophoblasts$celltype        <- Idents (Trophoblasts)

Idents (Trophoblasts) <- "celltype.model"
```

```{r, Supplemental figure 4A dotplot}
DefaultAssay (Trophoblasts) <- "RNA"
Idents       (Trophoblasts) <- "celltype.model"

my_levels <- c("CTB_MPI",    "CTB_TBpOrg",    "CTB_TBsOrg",
               "TSC_MPI",    "TSC_TBpOrg",    "TSC_TBsOrg",
               "cCTB_MPI",   "cCTB_TBpOrg",   "cCTB_TBsOrg",
               "EVT_MPI",    "EVT_TBpOrg",    "EVT_TBsOrg",
               "SCTp 1_MPI", "SCTp 1_TBpOrg", "SCTp 1_TBsOrg",
               "SCTp 2_MPI", "SCTp 2_TBpOrg", "SCTp 2_TBsOrg")

Trophoblasts@active.ident <- factor (Trophoblasts@active.ident, levels = my_levels)

DotPlot (Trophoblasts,
         features = c("ERVW-1",
                      "ERVFRD-1",
                      "ERVV-1",
                      "ERVV-2",
                      "ERV3-1",
                      "ERVH48-1",
                      "SLC1A5",
                      "MFSD2A",
                      "CGA",
                      "CGB",
                      "CGB1",
                      "CGB2",
                      "CGB5",
                      "CGB7",
                      "CGB8",
                      "INSL4",
                      "LEP",
                      "HSD17B1",
                      "PSG1",
                      "PSG2",
                      "PSG3",
                      "PSG4",
                      "PSG5",
                      "PSG6",
                      "PSG8",
                      "PSG9",
                      "PSG11",
                      "SDC1",
                      "KISS1",
                      "CCK",
                      "CSHL1",
                      "GH2",
                      "ANGPT2",
                      "INHBA"),
         assay     = "RNA",
         cols      = c("lightgrey", "red"),
         col.min   = 0,
         col.max   = 1,
         dot.min   = 0,
         dot.scale = 15) 
```

```{r, Supplemental figure 4B heatmap}
av.exp        <- AverageExpression (SCTp)$RNA
cor.exp       <- as.data.frame     (cor (av.exp))
cor.exp$x     <- rownames          (cor.exp)
cor.df        <- tidyr::gather     (data = cor.exp, y, correlation, c('SCTp 1', 'SCTp 2'))
cor.df

my_levels <- c("SCTp 1", "SCTp 2")
cor.df$x  <- factor (cor.df$x, levels = my_levels)
cor.df$y  <- factor (cor.df$y, levels = my_levels)

ggplot               (cor.df, aes (x, y, fill = correlation)) +
geom_tile            () +
scale_fill_gradientn (limits = c(0, 1), colours = colorRampPalette (rev (brewer.pal (n = 7, name = "RdYlBu"))) (100))
```

```{r, Supplemental figure 4C differential gene expression analysis}
DefaultAssay (SCTp) <- "RNA"

SCTp.response <- FindMarkers (SCTp,
                              ident.1         = c("SCTp 1"),
                              ident.2         = c("SCTp 2"),
                              test.use        = "MAST",
                              logfc.threshold = -Inf,
                              min.pct         = -Inf,
                              min.diff.pct    = -Inf,
                              verbose         = FALSE)
                     head    (SCTp.response, n = 100)
```

```{r}
write.xlsx (SCTp.response, file = "SCTp1v2_DGE.xlsx")
```

```{r, Supplemental figure 4C volcano plot}
pdf             ("SCTp1v2_DGE.pdf", width = 10, height = 10)
EnhancedVolcano (SCTp.response,
                 lab             = rownames (SCTp.response),
                 x               = "avg_log2FC",
                 y               = "p_val_adj",
                 FCcutoff        = 1.00,
                 pCutoff         = 10e-50,
                 pointSize       = 4.0,
                 col             = c('grey', 'grey', 'grey', 'red'),
                 colAlpha        = 1,
                 selectLab       = c('SLC40A1', 'PAGE4', 'KRT23', 'PEG10', 'VGLL1', 'ID2', 'MUC15', 'SLC26A2', 'ERVFRD-1', 'PSG6', 'PSG2', 'PSG3', 'PSG5', 'HIST1H1C', 'TFPI2', 'EBI3', 'ISG15', 'FLT1'),
                 xlim            = c(-6, 6),
                 ylim            = c(0, 325),
                 gridlines.major = FALSE,
                 drawConnectors  = TRUE, 
                 gridlines.minor = FALSE) + NoLegend () + ggtitle (NULL)
dev.off         ()
```
