---
title:  "MODEL-SPECIFIC DATA PREPARATION"
author: "Matthew J Shannon"
date:   "18/01/2024"
output: html_document
---

# Load in all dependencies

```{r, include = FALSE}
devtools::install_github ('chris-mcginnis-ucsf/DoubletFinder')
devtools::install_github ('satijalab/seurat-wrappers')
remotes::install_github  ("mojaveazure/seurat-disk")
devtools::install_github ("velocyto-team/velocyto.R")
BiocManager::install     ("pcaMethods")
BiocManager::install     ('MAST')
BiocManager::install     ('org.Hs.eg.db')
BiocManager::install     (version = "3.12")
BiocManager::install     ('Seurat')
BiocManager::install     ('readxl')
BiocManager::install     ('modes')
BiocManager::install     ('pheatmap')
BiocManager::install     ('limma')
BiocManager::install     ('clustree')
BiocManager::install     ('clusterProfiler')
BiocManager::install     ('EnhancedVolcano')
install.packages         ('fields')
install.packages         ('plotly')
install.packages         ("VennDiagram")
library                  (Seurat)
library                  (readr)
library                  (tidyverse)
library                  (Matrix)
library                  (dplyr)
library                  (ggplot2)
library                  (cowplot)
library                  (tibble)
library                  (readxl)
library                  (sctransform)
library                  (fields)
library                  (KernSmooth)
library                  (ROCR)
library                  (parallel)
library                  (reshape2)
library                  (pheatmap)
library                  (DoubletFinder)
library                  (limma)
library                  (SeuratWrappers)
library                  (SeuratDisk)
library                  (plotly)
library                  (clustree)
library                  (velocyto.R)
library                  (MAST)
library                  (EnhancedVolcano)
library                  (clusterProfiler)
library                  (AnnotationDbi)
library                  (org.Hs.eg.db)
library                  (VennDiagram)
library                  (RColorBrewer)
```

# Load the integrated trophoblast Seurat object

```{r}
load         (file = "~/Desktop/FINAL_Trophoblast_Data.Rdata")
DefaultAssay (Trophoblasts) <- "integrated"
```

```{r}
Trophoblasts$seurat_clusters <- Idents (Trophoblasts)
UMAPPlot (Trophoblasts, pt.size = 3, label = TRUE, split.by = 'Model') + NoLegend ()
```

# Subset the dataset into model-specific trophoblast groups

```{r}
# Maternal-placental interface (in vivo) trophoblasts
MPI <- subset (Trophoblasts, subset = Model == "MPI")
MPI

# PD-TOrg (in vitro) trophoblasts
TBpOrg <- subset (Trophoblasts, subset = Model == "TBpOrg")
TBpOrg

# hTSC-TOrg (in vitro) trophoblasts
TBsOrg <- subset (Trophoblasts, subset = Model == "TBsOrg")
TBsOrg
```

# Quantify the number of cells within each trophoblast state in each model

```{r, Table S5}
# Table S5B
table (Idents (MPI))

# Table S5C
table (Idents (TBpOrg))

# Table S5D
table (Idents (TBsOrg))
```

```{r}
rm (Trophoblasts)
```

# Prepare the maternal-placental interface (in vivo) trophoblast object

```{r}
DefaultAssay (MPI) <- "integrated"

#redo find neighbours/clusters
MPI <- FindNeighbors (MPI)
MPI <- FindClusters  (MPI, resolution = 0.375)
# Reduce trophoblast data dimensionality
MPI <- RunUMAP       (MPI, reduction  = "pca", dims = 1:30)
MPI <- RunTSNE       (MPI, reduction  = "pca", dims = 1:30)
MPI
```

```{r}
# Visualize the trophoblast data by TSNE
TSNEPlot    (MPI, label = TRUE, pt.size = 3) + NoLegend ()
# Visualize the trophoblast data by UMAP
UMAPPlot    (MPI, label = TRUE, pt.size = 3) + NoLegend ()
# Visualize expression of trophoblast QC metrics by UMAP
FeaturePlot (MPI, features = c("nFeature_RNA"), min.cutoff = 0, order = TRUE, pt.size = 3, cols = c("grey", "red"))
FeaturePlot (MPI, features = c("nCount_RNA"),   min.cutoff = 0, order = TRUE, pt.size = 3, cols = c("grey", "red"))
FeaturePlot (MPI, features = c("percent.mt"),   min.cutoff = 0, order = TRUE, pt.size = 3, cols = c("grey", "red"))
```

```{r}
# Bin states into trophoblast subtypes
MPI <- RenameIdents (MPI,
                     '0' = "CTB",
                     '1' = "CTB",
                     '2' = "CTB",
                     '3' = "cCTB",
                     '4' = "EVT",
                     '5' = "CTB",
                     '6' = "SCTp",
                     '7' = "cCTB")

MPI$celltype <- Idents (MPI)
MPI[[]]

UMAPPlot (MPI, label = TRUE) + NoLegend ()
```

# Identify enriched maternal-placental interface (in vivo) trophoblast-state gene markers

```{r}
DefaultAssay (MPI) <- "RNA"

MPI.markers <- FindAllMarkers (object   = MPI,
                               only.pos = TRUE,
                               min.pct  = 0.25, 
                               test.use = "MAST")

DefaultAssay (MPI) <- "integrated"
```

```{r}
MPI.markers %>% group_by (cluster) %>% top_n (20, avg_log2FC)
```

```{r, Table S3B}
write.xlsx (MPI.markers, file = "MPI_genemarkers.xlsx")
```

# Save the maternal-placental interface (in vivo) trophoblast object

```{r}
save (MPI, file = "MPI_Trophoblasts.Rdata")
```

# Demultiplex the TBpOrg (in vitro) trophoblast object

```{r}
Undifferentiated         <- read_tsv ("~/Desktop/BRC/PT_100/clusters.tsv")
Undifferentiated$barcode <- paste    ("47", Undifferentiated$barcode, sep = "_")
Differentiated           <- read_tsv ("~/Desktop/BRC/PT_100_EVT/clusters.tsv")
Differentiated$barcode   <- paste    ("48", Differentiated$barcode, sep = "_")
```

```{r}
TBpOrg@meta.data <- rownames_to_column (TBpOrg@meta.data , "barcode")
Demultiplex      <- rbind              (Undifferentiated, Differentiated)
```

```{r}
TBpOrg@meta.data <- merge              (TBpOrg[[]], Demultiplex, all = FALSE)
TBpOrg@meta.data <- column_to_rownames (TBpOrg@meta.data, "barcode")
```

```{r}
Idents (TBpOrg) <- "assignment"

TBpOrg <- subset (TBpOrg, idents = c("0", "1", "2"))
TBpOrg

Idents (TBpOrg) <- "celltype"
```

# Prepare the TBpOrg (in vitro) trophoblast object

```{r}
DefaultAssay (TBpOrg) <- "integrated"

#redo find neighbours/clusters
TBpOrg <- FindNeighbors (TBpOrg)
TBpOrg <- FindClusters  (TBpOrg, resolution = 0.15)
# Reduce trophoblast data dimensionality
TBpOrg <- RunUMAP       (TBpOrg, reduction  = "pca", dims = 1:30)
TBpOrg <- RunTSNE       (TBpOrg, reduction  = "pca", dims = 1:30)
TBpOrg
```

```{r}
# Visualize the trophoblast data by TSNE
TSNEPlot    (TBpOrg, label = TRUE, pt.size = 3) + NoLegend ()
# Visualize the trophoblast data by UMAP
UMAPPlot    (TBpOrg, label = TRUE, pt.size = 3) + NoLegend ()
# Visualize expression of trophoblast QC metrics by UMAP
FeaturePlot (TBpOrg, features = c("nFeature_RNA"), min.cutoff = 0, order = TRUE, pt.size = 3, cols = c("grey", "red"))
FeaturePlot (TBpOrg, features = c("nCount_RNA"),   min.cutoff = 0, order = TRUE, pt.size = 3, cols = c("grey", "red"))
FeaturePlot (TBpOrg, features = c("percent.mt"),   min.cutoff = 0, order = TRUE, pt.size = 3, cols = c("grey", "red"))
```

```{r}
# Bin states into trophoblast subtypes
TBpOrg <- RenameIdents (TBpOrg,
                        '0' = "CTB",
                        '1' = "cCTB",
                        '2' = "CTB",
                        '3' = "EVT",
                        '4' = "SCTp",
                        '5' = "TSC")

TBpOrg$celltype <- Idents (TBpOrg)
TBpOrg[[]]

UMAPPlot (TBpOrg, label = TRUE) + NoLegend ()
```

# Identify enriched PD-TOrg (in vitro) trophoblast-state gene markers

```{r}
DefaultAssay (TBpOrg) <- "RNA"

TBpOrg.markers <- FindAllMarkers (object   = TBpOrg,
                                   only.pos = TRUE,
                                   min.pct  = 0.25, 
                                   test.use = "MAST")

DefaultAssay (TBpOrg) <- "integrated"
```

```{r}
TBpOrg.markers %>% group_by (cluster) %>% top_n (20, avg_log2FC)
```

```{r, Table S3C}
write.xlsx (TBpOrg.markers, file = "TBpOrg_genemarkers.xlsx")
```

# Save the PD-TOrg (in vitro) trophoblast object

```{r}
save (TBpOrg, file = "TBpOrg.Rdata")
```

# Prepare the TBsOrg (in vitro) trophoblast object

```{r}
DefaultAssay (TBsOrg) <- "integrated"

#redo find neighbours/clusters
TBsOrg  <- FindNeighbors (TBsOrg)
TBsOrg  <- FindClusters  (TBsOrg, resolution = 0.25)
# Reduce trophoblast data dimensionality
TBsOrg  <- RunUMAP       (TBsOrg, reduction  = "pca", dims = 1:30)
TBsOrg  <- RunTSNE       (TBsOrg, reduction  = "pca", dims = 1:30)
TBsOrg
```

```{r}
# Visualize the trophoblast data by TSNE
TSNEPlot    (TBsOrg, label = TRUE, pt.size = 3) + NoLegend ()
# Visualize the trophoblast data by UMAP
UMAPPlot    (TBsOrg, label = TRUE, pt.size = 3) + NoLegend ()
# Visualize expression of trophoblast QC metrics by UMAP
FeaturePlot (TBsOrg, features = c("nFeature_RNA"), min.cutoff = 0, order = TRUE, pt.size = 3, cols = c("grey", "red"))
FeaturePlot (TBsOrg, features = c("nCount_RNA"),   min.cutoff = 0, order = TRUE, pt.size = 3, cols = c("grey", "red"))
FeaturePlot (TBsOrg, features = c("percent.mt"),   min.cutoff = 0, order = TRUE, pt.size = 3, cols = c("grey", "red"))
```

```{r}
# Bin states into trophoblast subtypes
TBsOrg <- RenameIdents (TBsOrg,
                        '0' = "TSC",
                        '1' = "cCTB",
                        '2' = "SCTp",
                        '3' = "CTB",
                        '4' = "SCTp",
                        '5' = "TSC",
                        '6' = "EVT" )

TBsOrg$celltype <- Idents (TBsOrg)
TBsOrg[[]]

UMAPPlot (TBsOrg, label = TRUE) + NoLegend ()
```

# Identify enriched hTSC-TOrg (in vitro) trophoblast-state gene markers

```{r}
DefaultAssay (TBsOrg) <- "RNA"

TBsOrg.markers <- FindAllMarkers (object   = TBsOrg,
                                     only.pos = TRUE,
                                     min.pct  = 0.25, 
                                     test.use = "MAST")

DefaultAssay (TBsOrg) <- "integrated"
```

```{r}
TBsOrg.markers %>% group_by (cluster) %>% top_n (20, avg_log2FC)
```

```{r, Table S3D}
write.xlsx (TBsOrg.markers, file = "TBsOrg_genemarkers.xlsx")
```

# Save the hTSC-TOrg (in vitro) trophoblast object

```{r}
save (TBsOrg, file = "TBsOrg.Rdata")
```